---
title: "Blog Post 1 - Validating Data and Priors"
output: html_document
---

```{r libraries, warning=FALSE,message=FALSE}
library(tidyverse)
library(data.table)
library(GGally)
library(ggrepel)
library(MASS)
rm(list=ls())
```


```{r load_data, warning=FALSE,message=FALSE}
calc_paul <- function(result, result_type, distance) {
  if(result_type == "seconds") {
    race_split <- 500 / (distance / result)
  } else if(result_type == "speed") {
    race_split <- 500 / result
  } else {
    warning("result type not recognised")
  }
  
  more_secs <- (log(distance/2000)/log(2)) * 5
  split_2k <- race_split - more_secs
  speed_2k <- 500 / split_2k
  return(speed_2k)
}

```

```{r}

cam_results <- readRDS("../data/clean_results/cam_race_results.rds") %>%
  setDT() %>%
  .[, paul_speed := calc_paul(seconds, "seconds", distance)] %>%
  .[!is.na(leg), race := paste(race, leg)] %>%
  .[, norm_ps := paul_speed / max(paul_speed), by = c("year", "gender", "race", "leg")]

bumps_data <- read.csv("../data/bumps.csv",
                       stringsAsFactors = FALSE) %>%
  rename_all(tolower) %>%
  dplyr::select(year, college, crew, gender, startpos) %>%
  setDT() %>%
  .[is.na(crew), crew := 1] %>%
  left_join(., cam_results, by = c("year", "college", "crew", "gender")) %>%
  filter(!is.na(norm_ps))

models <- bumps_data %>%
  dplyr::select(startpos, norm_ps, race, year, gender) %>%
  split(., f = list(.$race, .$year, .$gender)) %>%
  .[sapply(., nrow)>0] %>%
  lapply(., function(race_data) {
    model <- lm(norm_ps ~ startpos, data = race_data)
    fastest_crew_pred <- stats::predict.lm(model, newdata = data.frame(startpos = 1))
    df <- data.frame(race = unique(race_data$race),
                     year = unique(race_data$year),
                     gender = unique(race_data$gender),
                     multiplier = fastest_crew_pred)
  }) %>%
  do.call(rbind, .)

bumps_data <- bumps_data %>%
  left_join(., models, by = c("race", "year", "gender")) %>%
  mutate(racenorm_ps = norm_ps / multiplier)

p1 <- bumps_data %>%
  filter(startpos < 32) %>%
  ggplot(data = ., aes(x = startpos, y = racenorm_ps, colour = race)) +
  geom_point(alpha = 0.5) +
  stat_smooth(method = "lm") +
  #facet_grid(year ~ gender)
  facet_wrap(~gender)

p2 <- bumps_data %>%
  filter(startpos < 32) %>%
  ggplot(data = ., aes(x = startpos, y = racenorm_ps)) +
  geom_point(alpha = 0.5) +
  stat_smooth(method = "lm") +
  facet_wrap(~gender)


```

```{r}
p <- bumps_data %>%
    filter(year > 2015 & gender == "M" & startpos < 32) %>%
    ggplot(., aes(x = startpos, y = racenorm_ps)) +
    geom_point(alpha = 0.5) +
    stat_smooth(method = "lm") +
    theme_minimal()

model <- bumps_data %>%
  filter(year > 2015 & gender == "M" & startpos < 32) %>%
  lm(startpos ~ racenorm_ps, data = .)

predictions2 <- lapply(1:31, function(position) {
  prediction <- inverse.predict.lm(model, position)
  mean <- prediction$Prediction
  confidence <- prediction$Confidence
  df <- data.frame(position = position,
                   speed = mean,
                   conf = confidence)
}) %>%
  do.call(rbind, .) 

```


bumps_data <- read.csv("../data/bumps.csv",
                       stringsAsFactors = FALSE) %>%
  rename_all(tolower) %>%
  setDT() %>%
  .[is.na(crew), crew := 1] %>%
  melt.data.table(id.vars = c("competition", "college", "year", "crew", "gender", "startpos"),
                  value.name = "bump", variable.name = "day") %>%
  .[order(year, startpos, day)] %>%
  .[is.na(bump), bump := 0] %>%
  .[, campaign := cumsum(bump), by = c("competition", "college", "year", "crew", "gender")] %>%
  .[, dend_pos := startpos - campaign] %>%
  .[, dstart_pos := startpos - lag(campaign), by = c("competition", "college", "year", "crew", "gender")] %>%
  .[is.na(dstart_pos), dstart_pos := startpos] %>%
  left_join(., cam_results, by = c("year", "college", "crew", "gender")) %>%
  setDT() %>%
  .[order(year, competition, gender, day, dstart_pos)] %>%
  .[!is.na(paul_speed)]


p <- bumps_data %>%
  filter(day == "day4") %>%
  filter(year == 2018 & gender == "M") %>%
  ggplot(., aes(x = startpos, y = paul_speed, colour = race)) +
  geom_point(alpha = 0.5) +
  stat_smooth(method = "lm") +
  theme_minimal() +
  xlab("assumed_speed_order") +
  ylab("speed")
```

```{r}
upperfun <- function(data,mapping){
  ggplot(data = data, mapping = mapping)+
    geom_density2d()+
    scale_x_continuous(limits = c(-1,1))+
    scale_y_continuous(limits = c(-1,1))
}   

lowerfun <- function(data,mapping){
  ggplot(data = data, mapping = mapping)+
    geom_point(alpha = 0.5)+
    scale_x_continuous(limits = c(-1,1))+
    scale_y_continuous(limits = c(-1,1))
}  

#need to remove the duplicates in cam race results
self_correlate <- cam_results %>%
  .[, mean_diff := paul_speed - mean(paul_speed), by = c("year", "race", "gender")] %>%
  .[, c("year", "college", "crew", "gender", "race", "mean_diff")] %>%
  dcast(year + college + crew + gender ~ race, value.var = "mean_diff") %>%
  ggpairs(., mapping=ggplot2::aes(colour = factor(year)),
          columns = 5:ncol(.),
          upper = list(continuous = wrap(upperfun)),
          lower = list(continuous = wrap(lowerfun)))
  
```

```{r}
darwin <- cam_results %>%
  filter(year == 2019 & college == "Darwin" & crew == 1 & gender == "M") %>%
  mutate(startpos = 29)

p <- bumps_data %>%
  filter(year > 2015 & gender == "M" & race %in% c("Newnham Short Course", "Fairbairns") & day == "day4") %>%
  ggplot(., aes(x = startpos, y = norm_ps)) +
  geom_point(aes(colour = year), alpha = 0.2) +
  geom_point(data = filter(bumps_data, year > 2015 & gender == "M" & race %in% c("Newnham Short Course", "Fairbairns") & day == "day4" & campaign >= 2), aes(x = startpos, y = paul_speed, colour = year)) +
  scale_colour_gradient(low = "darkred", high = "darkblue") +
  geom_point(data = darwin, shape = 21, fill = "blue", colour = "red", size = 2) +
  geom_text_repel(data = darwin, label = "darwin m1 2019", colour = "darkblue") +
  #facet_grid(gender~year) +
  #facet_wrap(year~race) +
  #facet_wrap(~gender) +
  xlab("Bumps Starting Position") +
  ylab("Paul's Law Adjust Speed (m/s)") +
  ggtitle("historic bumps starting position against cam race speeds for Lent crews",
          subtitle = "data 2016-2018, race speeds adjusted via Paul's law") +
  stat_smooth(method = "lm") +
  theme_minimal()

```

```{r}
model <- bumps_data %>%

```

